Exceedance probability comparison for the simplified Naomi model

### 2023-06-22 src/docs_paper/paper.Rmd
Replace all elgm-inf with naomi-aghq
Upload paper and appendix new versions #55

Naomi is a complex model, and as such presents a challenging Bayesian inference problem.
As well as hundreds of latent field parameters, Naomi has >20 hyperparameters: substantially more than the small number that can typically be handled by integrated nested Laplace approximations [INLA; @rue2009approximate].
Moreover, observations depend on multiple structured additive predictors, such that Naomi falls into the class of extended latent Gaussian models [ELGMs; @stringer2022fast].
As such, the model is not easily specified by any software which requires a formula interface.

@@

We fit the simplified Naomi model (Section \ref{sec:naomi}) to data from Malawi using three inferential approaches.
The three approaches were:
1. TMB,
2. PCA-AGHQ, and
3. NUTS: the Hamiltonian Monte Carlo (HMC) algorithm No-U-Turn Sampling (NUTS) using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
The `TMB` C++ user-template used to specify the log-posterior was the same for each.
The dimension of the latent field was $N = 467$ and the dimension of the hyperparameters was $m = 24$.
Settings used for each inferential method are provided in Table \ref{tab:inference-methods}, and, where relevant, discussed further below.
For the deterministic methods, following inference we simulated hyperparameter and latent field samples.
For all methods, we simulated age-sex-district specific HIV prevalence, ART coverage and HIV incidence from the latent field and hyperparameter posteriors.
<!-- Note: eventually this figure should be generated using the algorithm proposed in the paper -->
Example model outputs from TMB are illustrated in Figure \ref{fig:naomi-results}.
The \textsc{R} [@r] code used to produce all results we describe below is available at `github.com/athowes/elgm-inf`.
We used `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].

@@

```{r node-positions, fig.cap="The 6561 PCA-AGHQ node positions (green, rug plot) projected onto the hyperparameter marginal posteriors (grey, histogram) for each of the 24 hyperparameters. Some hyperparameters, such as \\texttt{logit_phi_alpha_x}, are well covered, where as others, such as \\texttt{log_sigma_lambda_x}, are near only being covered by one unique node."}
knitr::include_graphics("depends/nodes-samples-comparison.png")

@@

(ref:mean-sd) Barplot displaying frequency of dependencies for all of the R packages in the CRAN repository to date `r Sys.Date()

```{r mean-sd, fig.cap="."}
knitr::include_graphics("depends/mean-sd-alt.png")
```

>>>>

Naomi is a complex model, and as such presents a challenging Bayesian inference problem.
As well as hundreds of latent field parameters, Naomi has >20 hyperparameters: substantially more than the small number that can typically be handled by integrated nested Laplace approximations [INLA; @rue2009approximate].
Moreover, observations depend on multiple structured additive predictors, such that Naomi falls into the class of extended latent Gaussian models [ELGMs; @stringer2022fast].

@@

We fit the simplified Naomi model (Section \ref{sec:naomi}) to data from Malawi using three inferential approaches.
The three approaches were:
1. TMB,
2. PCA-AGHQ, and
3. NUTS: the Hamiltonian Monte Carlo (HMC) algorithm No-U-Turn Sampling (NUTS) using Stan [@carpenter2017stan] via the `tmbstan` package [@monnahan2018no].
The `TMB` C++ user-template used to specify the log-posterior was the same for each.
The dimension of the latent field was $N = 467$ and the dimension of the hyperparameters was $m = 24$.
Settings used for each inferential method are provided in Table \ref{tab:inference-methods}, and, where relevant, discussed further below.
For the deterministic methods, following inference we simulated hyperparameter and latent field samples.
For all methods, we simulated age-sex-district specific HIV prevalence, ART coverage and HIV incidence from the latent field and hyperparameter posteriors.
<!-- Note: eventually this figure should be generated using the algorithm proposed in the paper -->
Example model outputs from TMB are illustrated in Figure \ref{fig:naomi-results}.
The \textsc{R} [@r] code used to produce all results we describe below is available at `github.com/athowes/naomi-aghq`.
We used `orderly` [@orderly] for reproducible research, `ggplot2` for data visualisation [@wickham2016ggplot2] and `rticles` [@allaire2022rticles] for reporting via `rmarkdown` [@allaire2022rmarkdown].

@@

```{r node-positions, fig.cap="The 6561 PCA-AGHQ node positions (green, rug plot) projected onto the hyperparameter marginal posteriors (grey, histogram) for each of the 24 hyperparameters. Some hyperparameters, such as logit\\_phi\\_alpha\\_x, are well covered where as others, such as log\\_sigma\\_lambda\\_x, are near only being covered by one unique node."}
knitr::include_graphics("depends/nodes-samples-comparison.png")

@@

```{r mean-sd, fig.cap="PCA-AGHQ modestly improves estimation of the posterior mean, and substantially improves estimation of the posterior standard deviation, as compared with TMB."}
knitr::include_graphics("depends/mean-sd-alt.png")
```

### 2023-06-22 src/naomi-simple_exceedance/exceedance.Rmd
Upload paper and appendix new versions #55

exceedance_manuscript_plot <- df_wide %>%
  pivot_longer(cols = c("TMB", "aghq"), names_to = "method", values_to = "estimate") %>%
  mutate(
    method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq"),
    indicator = fct_relevel(indicator, "Second 90", "High incidence")
  ) %>%
  ggplot(aes(x = tmbstan, y = estimate)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    coord_fixed(ratio = 1) +
    scale_colour_manual(values = c("#56B4E9", "#999999"), labels = c("Female", "Male")) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed") +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "", col = "Sex") +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )

exceedance_manuscript_plot <- exceedance_manuscript_plot +
  geom_text(
    data = df_metrics %>% mutate(
      method = fct_relevel(method, "TMB", "PCA-AGHQ"),
      indicator = fct_relevel(indicator, "Second 90", "High incidence")    
    ),
    aes(x = -Inf, y = Inf, label = label),
    size = 3, hjust = 0, vjust = 1.5
  )

ggsave("exceedance.png", exceedance_manuscript_plot, h = 6.5, w = 6.25)

>>>>

exceedance_manuscript_plot <- df_wide %>%
  pivot_longer(cols = c("TMB", "aghq"), names_to = "method", values_to = "estimate") %>%
  mutate(
    method = fct_reorg(method, "TMB" = "TMB", "PCA-AGHQ" = "aghq"),
    indicator = fct_relevel(indicator, "Second 90", "High incidence")
  ) %>%
  ggplot(aes(x = tmbstan, y = estimate - tmbstan)) +
    geom_point(aes(col = sex), shape = 1, alpha = 0.6) +
    scale_colour_manual(values = c("#56B4E9", "#999999"), labels = c("Female", "Male")) +
    geom_abline(slope = 0, intercept = 0, linetype = "dashed", size = 0.25) +
    facet_grid(indicator ~ method) +
    labs(x = "NUTS", y = "Approximate - NUTS", col = "Sex") +
    geom_text(
      data = df_metrics %>% mutate(
        method = fct_relevel(method, "TMB", "PCA-AGHQ"),
        indicator = fct_relevel(indicator, "Second 90", "High incidence")    
      ),
      aes(x = -Inf, y = Inf, label = label),
      size = 3, hjust = 0, vjust = 1.5
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )

ggsave("exceedance.png", exceedance_manuscript_plot, h = 6.5, w = 6.25, bg = "white")
